What to simulate? Inferring the right direction for mental rotation

### 2014-02-05 man/cogsci-2014/cogsci-2014.tex
Update paper with new figure

% \begin{figure}[t]
%   \centering
%   \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_accuracy.pdf}}%
%   \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_time.pdf}}
%   \caption{\textbf{Effect of learning.} There was a clear effect of
%     learning on participants' responses, especially in the first half
%     of the experiment. Left: accuracy as a function of trial. Right:
%     response time in milliseconds as a function of trial.}
%   \label{fig:learning}
% \end{figure}

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{../../figures/D/theta_time_corrs.pdf}
    \caption{\textbf{Angle vs. response time}}
    \label{fig:theta-time-corrs}
  \end{subfigure}%
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{../../figures/D/theta_accuracy_corrs.pdf}
    \caption{\textbf{Angle vs. accuracy}}
    \label{fig:theta-accuracy-corrs}
  \end{subfigure}%
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{../../figures/D/human_model_corrs.pdf}
    \caption{\textbf{Human vs. model}}
    \label{fig:human-model-corrs}
  \end{subfigure}
  \vspace{10pt}
  \caption{\textbf{Correlations.} (a) Spearman correlations between
    the minimum angle of rotation and response time. (b) Spearman
    correlations between the minimum angle of rotation and
    accuracy. The \Oc{} and \Th{} models had 100\% accuracy;
    therefore, their correlations are undefined. (c) Pearson
    correlations between human and model (except the first set of
    bars, which is human vs. human), both for response time and
    accuracy. All error bars are 95\% confidence intervals.}
\end{figure*}

@@

For analyses of response time, confidence intervals around harmonic
means were computed using a bootstrap analysis using 10000 bootstrap
samples (sampled with replacement). Response times were only
considered for correct responses.  We also used a bootstrap analysis
of 10000 bootstrap samples to compute the confidence intervals around
both Spearman ($r_s$) and Pearson ($r$) correlations.  Unless
otherwise specified, all correlations were computed over 720
datapoints (one for each stimulus). For analyses of accuracy,
confidence intervals were computed from a binomial proportion with a
Jeffrey's beta prior.  To test if participants or a model was above
chance on a particular stimulus, we used the same binomial proportion,
and simply tested whether $p>0.05$ for the 50\% mark.

\TODO{need to recompute correlations}

@@

The average response time across all stimuli was \ExpTime{}
milliseconds.  As illustrated in Fig.~\ref{fig:theta-time-corrs}, the
minimum angle of rotation was significantly monotonically correlated
with average per-stimulus response times, both for ``flipped'' and
``same'' pairs. While this replicates the general result of previous
experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are
not as perfectly linear. However, \citeA{Gardony:2013gn} found a
constant effect for ``flipped'' pairs; an average of their results and
the classic results would produce something similar to ours.

The average accuracy across all stimuli was \ExpAccuracy{}, though
there were \ExpNumChance{} stimuli (out of 720) for which people were
not above chance.  The minimum angle was also correlated with
participants' average per-stimulus accuracy, as shown in
Fig.~\ref{fig:theta-accuracy-corrs}, though significantly more so for
``same'' pairs. This is the same result found both by
\citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.

To gauge between-subject reliability, we randomly split subjects into
two groups, computed the average response time and accuracy per
stimulus for each group, and then computed correlations for these two
measures. We repeated this procedure 100 times, and from those samples
computed the median correlations and 95\% confidence intervals. For
response time, the between-subjects correlation was \ExpTimeCorr{},
and for accuracy, it was \ExpAccuracyCorr{} (also shown in
Fig.~\ref{fig:human-model-corrs}).

@@

As we would expect, the number of actions taken by the \Oc{} model was
perfectly correlated with the minimum angle of rotation
(Fig.~\ref{fig:theta-time-corrs}), both for ``same'' and ``flipped''
pairs. Unsurprisingly, the \Oc{} model was the best fit to human
response times, with a correlation of \OcTimeCorr{}
(Fig.~\ref{fig:human-model-corrs}).  However, the \Oc{} model was
100\% accurate, and therefore the correlation between minimum angle
and accuracy was undefined (Fig.~\ref{fig:theta-accuracy-corrs}).

@@

The number of actions taken by the \Th{} model was significantly
correlated with the minimum angle (Fig.~\ref{fig:theta-time-corrs}),
though this effect was less strong for individual shapes (e.g.,
Fig.~\ref{fig:response-time-stimulus}).  The \Th{} model was able to
explain a moderate amount of the variance in human response times,
with a correlation of \ThTimeCorr{}
(Fig.~\ref{fig:human-model-corrs}). Like the \Oc{} model, the \Th{}
model had 100\% accuracy, and thus there was no correlation between
minimum angle and accuracy.

@@

The \Hc{} was the only model for which there was no effect of rotation
on response time (Fig.~\ref{fig:theta-time-corrs}). Although there was
a significant effect of rotation on accuracy
(Fig.~\ref{fig:theta-accuracy-corrs}), the \Hc{} model was barely
above chance (\HcAccuracy{}) and there were \HcNumChance{} individual
stimuli for which it was not above chance. The \Hc{} model was not a
particularly good predictor of human behavior, neither for response
times (\HcTimeCorr{}) nor accuracy (\HcAccuracyCorr{}), as shown in
Fig.~\ref{fig:human-model-corrs}.

@@

As with most of the other models, there was a significant effect of
rotation on the number of steps taken by the \Bq{} model
(Fig.~\ref{fig:theta-time-corrs}). The \Bq{} model explained variance
in human response times about as well as the \Th{} model
(Fig.~\ref{fig:human-model-corrs}), with a correlation of
\BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
prior.

The \Bq{} model was quite accurate overall (equal prior:
(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
there was \BqNumChance{} stimulus for which it was not above chance;
with the unequal prior, there were \BqpNumChance{}. There was a
significant effect of rotation on the \Bq{} model's accuracy
(Fig.~\ref{fig:theta-accuracy-corrs}), and it explained about the same
amount of variance in people's accuracy as the \Hc{} model (equal
prior: \BqAccuracyCorr{}; unequal prior: \BqpAccuracyCorr{}).

>>>>

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
    \caption{\textbf{Model vs. human response times.}}
    \label{fig:human-model-scatters}
  \end{center}
\end{figure*}

@@

For analyses of response time, confidence intervals around harmonic
means were computed using a bootstrap analysis using 10000 bootstrap
samples (sampled with replacement). Response times were only
considered for correct responses.  We also used a bootstrap analysis
of 10000 bootstrap samples to compute the confidence intervals around
both Spearman ($r_s$) and Pearson ($\rho$) correlations.  Unless
otherwise specified, all correlations were computed over 720
datapoints (one for each stimulus). For analyses of accuracy,
confidence intervals were computed from a binomial proportion with a
Jeffrey's beta prior.  To test if participants or a model was above
chance on a particular stimulus, we used the same binomial proportion,
and simply tested whether $p>0.05$ for the 50\% mark.

@@

The average response time across all stimuli was \ExpTime{}
milliseconds.  The minimum angle of rotation was significantly
monotonically correlated with average per-stimulus response times,
both for ``flipped'' and ``same'' pairs \TODO{}. While this replicates
the general result of previous experiments
\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as
perfectly linear. However, \citeA{Gardony:2013gn} found a constant
effect for ``flipped'' pairs; an average of their results and the
classic results would produce something similar to ours.

The average accuracy across all stimuli was \ExpAccuracy{}, though
there were \ExpNumChance{} stimuli (out of 720) for which people were
not above chance.  The minimum angle was also correlated with
participants' average per-stimulus accuracy, though significantly more
so for ``same'' pairs (\TODO{}). This is the same result found both by
\citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.

To gauge between-subject reliability, we randomly split subjects into
two groups, computed the average response time and accuracy per
stimulus for each group, and then computed correlations for these two
measures. We repeated this procedure 100 times, and from those samples
computed the median correlations and 95\% confidence intervals. For
response time, the between-subjects correlation was \ExpTimeCorr{},
and for accuracy, it was \ExpAccuracyCorr{}.

@@

As we would expect, the number of actions taken by the \Oc{} model was
perfectly correlated with the minimum angle of rotation
(Fig.~\ref{fig:response-time-accuracy}). Unsurprisingly, the \Oc{}
model was the best fit to human response times, with a correlation of
\OcTimeCorr{} (Fig.~\ref{fig:human-model-scatters}).  However, the
\Oc{} model was 100\% accurate, and therefore the correlation between
minimum angle and accuracy was undefined.

@@

There was an overall monotonic relationship between the minimum angle
of rotation and the number of actions taken by the \Th{} model
(Fig.~\ref{fig:response-time-accuracy}). However, the relationship
between angle and response time for individual shapes was usually not
monotonic (e.g., Fig.~\ref{fig:response-time-stimulus}).  The \Th{}
model was able to explain a moderate amount of the variance in human
response times, with a correlation of \ThTimeCorr{}
(Fig.~\ref{fig:human-model-scatters}). Like the \Oc{} model, the \Th{}
model had 100\% accuracy, and thus there was no correlation between
minimum angle and accuracy.

@@

The \Hc{} was the only model for which there was no monotonic
relationship between rotation and response time
(Fig.~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model was
barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
individual stimuli for which it was not above chance. The \Hc{} model
was not a particularly good predictor of human response times
(\HcTimeCorr{}), as shown in Fig.~\ref{fig:human-model-scatters}. It
was a moderately good predictor of human accuracy (\HcAccuracyCorr{}).

@@

Like the \Oc{} and \Th{} models, there was an overall monotonic
relationship between rotation and the number of steps taken by the
\Bq{} model (Fig.~\ref{fig:response-time-accuracy}). Unlike the \Th{}
model, this relationship existed for individual shapes as well (e.g.,
Fig.~\ref{fig:response-time-stimulus}).  The \Bq{} model explained
variance in human response times about as well as the \Th{} model
(Fig.~\ref{fig:human-model-scatters}), with a correlation of
\BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
prior.

The \Bq{} model was quite accurate overall (equal prior:
(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
there was \BqNumChance{} stimulus for which it was not above chance;
with the unequal prior, there were \BqpNumChance{}. The \Bq{} model
explained about the same amount of variance in people's accuracy as
the \Hc{} model (equal prior: \BqAccuracyCorr{}; unequal prior:
\BqpAccuracyCorr{}).
